### 2022-10-06 submission/manuscript.tex
Bring in Table 1 from Singletons papper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\raggedright
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{itemize}
\tightlist
\item
  What is rarefaction? History, reason for rarefaction

@@

\hypertarget{datasets}{%
\subsubsection{Datasets}\label{datasets}}

\begin{itemize}
\tightlist
\item
  Represented diverse communities
\item
  High quality V4 sequence data that consisted of paired 250 nt reads
\item
  Wide distribution in the number of reads within each dataset (Figure
  S1, Table 1)
\end{itemize}

\hypertarget{null-models}{%
\subsubsection{Null models}\label{null-models}}

\begin{itemize}
\item
  Null models were samplings, without replacement, of all samples from a
  study that have been pooled together and sampled to the same depth of
  sampling
\item
  Created 100 random sets of samples per dataset
\item
  Would expect a zero correlation between alpha diversity metric and
  number of sequences in the community
\item
  Rarefaction was the only approach that resulted in a near-zero
  correlation (Figure 1)
\item
  Would expect a zero correlation between beta diversity metrics and the
  difference in the number of sequences in the samples going into the
  distance calculation
\item
  Rarefaction was the only approach that resulted in a near-zero
  correlation (Figure 2)
\end{itemize}

\hypertarget{result-when-sampling-effort-is-confounded-with-treatment-group}{%
\subsubsection{Result when sampling effort is confounded with treatment
group}\label{result-when-sampling-effort-is-confounded-with-treatment-group}}

\begin{itemize}
\item
  Used Null model
\item
  Created two treatment groups for each study by dividing samples based
  on whether they were above or below the median number of sequences per
  sample
\item
  Used Wilcoxon test to detect differences between treatment groups for
  alpha diversity, would expect \textasciitilde5\% of random sets to be
  significant
\item
  Rarefaction the only approach that reliably resulted in the expected
  Type I error rate
\item
  Used AMOVA test to detect differences between treatment groups for
  beta diversity, would expect \textasciitilde5\% of random sets to be
  significant
\item
  Rarefaction the only approach that reliably resulted in the expected
  Type I error rate
\end{itemize}

\hypertarget{result-when-known-effect-size-imposed-between-two-groups-of-samples}{%
\subsubsection{Result when known effect size imposed between two groups
of
samples}\label{result-when-known-effect-size-imposed-between-two-groups-of-samples}}

\begin{itemize}
\item
  For each dataset I divided samples into two groups. The probability
  distribution for the first group was determined by the null model and
  the second was determined by xxxxxxxxxxx
\item
  These parameters were selected because they produced results were some
  of the simulations for the rarefied data did not result in a
  significant test
\item
  Fraction of tests that were significant would be a measure of
  statistical power
\item
  Used Wilcoxon test to detect differences between treatment groups for
  alpha diversity
\item
  All tests resulted in reduced power to detect differences relative to
  the power of using rarefaction
\item
  Used AMOVA test to detect differences between treatment groups for
  beta diversity
\item
  All tests resulted in reduced power to detect differences relative to
  the power of using rarefaction
\item
  This simulation would work well for measures based on the structure of
  the community, but lacked resolution for those based on the richness
  or membership of commmunities. For an additional simulation, the first
  treatment group followed a null distribution for the second, XX\% of
  OTUs were removed from the distribution
\item
  Tests for richness and membership (i.e.~Jaccard) resutled in reduced
  power relative to rarefaction
\end{itemize}

\hypertarget{effect-of-sampling-depth-on-metrics}{%
\subsubsection{Effect of sampling depth on
metrics}\label{effect-of-sampling-depth-on-metrics}}

\begin{itemize}
\tightlist
\item
  Simulated rarefying null model datasets to retain varying percentage
  of samples
\item
  High sampling depth resulted in less variation between samples for
  alpha and beta diversity
\end{itemize}

\hypertarget{re-analysis-of-previously-published-datasets}{%
\subsubsection{Re-analysis of previously published
datasets}\label{re-analysis-of-previously-published-datasets}}

@@

@@

  \begin{itemize}
  \tightlist
  \item
    Many recommended methods are borrowed from gene expression analysis
  \item
    Meaning of zeroes in data - structural vs.~below limit of detection
  \end{itemize}
\item
  Factors that determine what number of sequences to rarefy to
\item
  Need better methods of pooling libraries that result in more even
  distribution of sequences across samples
\item
  Rarefy your data
\end{itemize}

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}

\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1.0in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage[shortcuts]{extdash}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

\begin{itemize}
\tightlist
\item
  Motivation

  \begin{itemize}
  \tightlist
  \item
    Problem of uneven sampling effort
  \end{itemize}
\item
  What is rarefaction? History, reason for rarefaction

@@

\textbf{\emph{Choice of datasets.}} I selected 16S rRNA gene sequence
data from from 12 studies that characterized the variation in bacterial
communities from diverse environments (Table 1). The specific studies
were selected because their data was publicly accessible through the
Sequence Read Archive, the original investigators sequenced the V4
region of the 16S rRNA gene using paired 250 nt reads, and my previous
familiarity with the data. The use of paired 250 nt reads to sequence
the V4 region resulted in a near complete two-fold overap of the V4
region resulting in high quality contigs with a low sequencing error
rate (1). These data were processed through the standard sequence
curation pipeline to generate operational taxonomic units (OTUs) using
the mothur software package (1, 2). The original studies generated the
sequence data by pooling separate PCR products that were generated by
amplifying the V4 region of the 16S rRNA gene from the bacterial DNA in
multiple samples. Because pooling equimolar quantities of DNA is frought
with difficulties, it was common to observe wide variation in the number
of sequences in each sample (Figure S1).

\textbf{\emph{Without rarefaction, metrics of alpha diversity are
sensitive to sampling effort.}} To test the sesitivity of various
approaches of measuring alpha diversity to sampling effort, I generated
null models for each study. Under a null model, each sample from the
same study would be expected to have the same alpha diversity regardless
of the sampling effort. I assessed richness without any correction,
using normalized OTU counts, with estimates based on non-parametric and
parametric approaches, and using rarefaction. For each study, all of the
approaches, except for rarefaction, showed a strong correlation between
richness and the number of sequences in the sample (Figure 1A). Next, I
assessed diversity using the Shannon diversity index and the inverse
Simpson diversity index without any correction, using normalized OTU
counts, and rarefaction; I also used a non-parametric estimator of
Shannon diversity. The correlation between sampling depth and the
diversity metric was not as strong as it was for richness and the
inverse Simpson diversity values were less sensitive than the Shannon
diversity values; however, the correlation to the rarefied diversity
metrics were the lowest for all of the metrics and studies (Figure 1B).
The rarefied alpha-diversity metrics consistently demonstrated a lack of
sensitivity to sampling depth.

\textbf{\emph{Without rarefaction, metrics of beta diversity are
sensitive to sampling effort.}} To test the sesitivity of various
approaches of measuring beta diversity to sampling effort, I used the
same null models used for studying the sensitivity of alpha diversity.
Under a null model, the ecological distance between any pair of samples
would be the same regardless of the difference in the number of
sequences observed in each sample. First, I analyzed the sensitivity of
the Jaccard distance coefficient, which incorporates whether an OTU is
present in each community and not their relative abundance. When
calculating Jaccard distances using the uncorrected OTU counts,
normalized OTU counts, relative abundances, and rarefaction only the
rarefied data showed a lack of sensitivity to sampling effort (Figure
2A). Second, I analyzed the sensitivity of the Bray-Curtis distance
coefficient, which is a popular metric that incorporates the abundance
of each OTU. Similar to what I observed with the Jaccard coefficient,
only the rarefied data showed a lack of sensitivity to sampling effort
(Figure 2B). Third, I calcualted Aitchison distances on raw OTU counts
where the central log-ratio (CLR) was calculated by ignoring OTUs in
samples with zero counts (robust CLR), adding a pseudocount of 1 to all
OTU counts prior to calculating the CLR (one CLR), XXXXX XXXXX XXXXX
XXXXX XXXXX XXXXX XXXXX XXXXX XXXXX XXXXX XXXXXXXXXX (n CLR), and
imputing the value of zero counts (z CLR). Regardless of the approach,
the Aitchison distances were all strongly sensitive to sampling effort
(Figure 2C). Finally, I used the cumulative sum scaling (CSS)
normalization from metagenomeSeq and variance stabilization technique
(VST) from DeSeq2 prior to calculating Euclidean distances. Both
approachs revealed a strong sensitivity to sampling effort (Figure 2D).
Although Euclidean distances are not typically used on raw or rarefied
count data in ecology, rarefied Euclidean distances were not sensitive
to sampling effort. Across each of the beta diversity metrics and
approaches used to account for uneven sampling effort and sparsity,
rarefaction was the least sensitive approach to differences in sampling
effort.

\textbf{\emph{Rarefaction limits the detection of false positives when
sampling effort and treatment group are confounded.}} Next, I
investigated the impact of the various strategies and metrics on falsely
detecting a significant diffeerence using the the same communities
generated from the null model in the analysis of alpha and beta
diversity metrics. To test for differences in alpha and beta diversity I
used the non-parametric Wilcoxon test and non-parametric
permutation-based multivariate analysis of variance (PERMANOVA). First,
within each study, I randomly assigned each sample to one of two
treatment groups. My expectation was that approximately 5 of the 100
(5\%) random tests for each comparison would yield a significant test
result. Indeed, for each study and alpha and beta diversity metric and
strategy for accounting for uneven sampling, approximately 5\% of the
tests yielded a significant result (Figure 3). Second, within each
study, I assigned samples with more than the median number of sequences
per sample to one treatment group and the rest to another treatment
group. If there is no sensitivity to sampling effort, I would again
expect that 5\% of the tests would yield a significant result. In fact,
only the rarefied data consistently resulted in a 5\% false positive
rate for alpha and beta diversity metrics (Figure 4). These results
align with the observed sensitivity of alpha and beta diversity metrics
to sampling effort and underscore the value of rarefaction.

\textbf{\emph{Rarefaction preserves the statistical power to detect
differences between treatment groups.}} To assess the impact of
different approaches to control for uneven sampling effort I performed
two additional simulations. In the first simulation, for each study
samples were randomly assigned to one of two treatment groups. Samples
in the first treatment group were generated by sampling from the null
distribution. Samples in the second treatment group were generated by
perturbing the null distribution by increasing the relative abundance of
10\% of the OTUs by 5\%. These values were determined after empirically
searching for conditions that resulted in a large fraction of the
randomizations yieleding a significant result across most of the
studies. The fraction of tests that yielded a significant test was a
measure of the statistical power for the test. Relative to the rarefied
data, the power to detect differences in alpha and beta diversity was
considerably lower for each of the strategies for handling uneven
sampling effort. The power to detect differences in richness by all
approaches was low (Figures 7 and 8). This was likely because the
approach to generating the second community did not necessarily change
the number of OTUs in each treatment group. To explore this further, in
the second simulation the second treatment group was perturbed by
removing 3\% of the OTUs from the model. Again, the rarefied data had a
considerably higher statistical power than the other approaches when
measuring richness (Figure 9). Both simulations highlight the value of
rarefaction for preserving the statistical power to detect differences
between treatment groups for measures of alpha and beta diversity.

\textbf{\emph{Increased rarefaction depth reduces intra-sample variation
in alpha and beta diversity.}} To assess the sensitivity of alpha and
beta diversity metrics to rarefaction depth, I again used the dataset
generated using the null models and rarefied them to varying depths. For
the alpha diversity metrics, the value of the metrics plateaued and the
coefficient of variation (i.e., the mean divided by the standard
deviation) between samples remained constant as the rarefaction depth
increased (Figure 10). For beta diversity metrics, the distance between
samples and the coefficient of variation between samples decreased as
sampling depth increased (Figure 11). These results confirm that greater
sequencing depth provides a more robust estimate of the metrics.

\textbf{\emph{Most studies have a high level of sequencing coverage.}} I
calculated the Good's coverage for the observed data and found that each
of the studies had a minimum coverage greater than 90\% at their lowest
sequencing depth (Figure 12A). This suggests that most studies have a
high level of sequencing coverage. Next, I returned to the null models
to ask how much sequencing effort was required to obtain higher levels
of coverage. To obtain 95 and 99\% coverage, an average of XX and
XX-fold more sequence data was required, respectively (Figure 12B).
Although it is clear that most researchers would desire greater
coverage, the sequencing effort required to acheive that sequencing
depth would likely limit the number of samples that could be sequenced
when controlling for costs.

\textbf{\emph{Evidence that not rarefying data can impact results in
original studies.}} To assess the impact of using alternatives to
rarefaction, I reassessed hypotheses from the human, mice, and
peromyscus studies. These studies were selected because they originated
from my research group and I know the datasets well. First, the study
that published the human dataset was interested in the difference in the
fecal microbial communities of patients with and without colorectal
cancer. Here, I divided the patients into peope with and without screen
relevant neoplasia (SRN). Among the alpha diversity metrics, the test
were non-significant with similar effect sizes; however, the richness
estimates obtained using parametric appraoches detected statistically
significant differences in richness. Although the tests using
Bray-Curtis distances were significant regardless of whether rarefaction
was used, the effect sizes were smaller for the distances calculated
using the CLR, VST, CSS transformations.

@@

Mouse

Peromyscus

@@

  \begin{itemize}
  \tightlist
  \item
    Many recommended methods are borrowed from gened expression analysis
  \item
    Meaning of zeroes in data - structural vs.~below limit of detection
  \end{itemize}
\item
  Factors that determine what number of sequences to rarefy to
\item
  Need better methods of pooling libraries that result in more even
  distribution of sequences across samples
\item
  Rarefy your data
\end{itemize}

@@

\textbf{Null models.} Null models were generated by randomly assigning
each sequence in the study to an OTU and sample while keeping constant
the number of sequences per sample and the total number of sequences in
each OTU. Because the construction of the null models was a stochastic
process, 100 replicates were geneated for each study.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{0}{1}
\leavevmode\vadjust pre{\hypertarget{ref-Kozich2013}{}}%
\CSLLeftMargin{1. }%
\CSLRightInline{\textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter
NT}, \textbf{Highlander SK}, \textbf{Schloss PD}. 2013. {Development of
a dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform}.
Applied and environmental microbiology \textbf{79}:5112--5120.}

\leavevmode\vadjust pre{\hypertarget{ref-Schloss2009}{}}%
\CSLLeftMargin{2. }%
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.}

\leavevmode\vadjust pre{\hypertarget{ref-Li2015}{}}%
\CSLLeftMargin{3. }%
\CSLRightInline{\textbf{Li Q}, \textbf{Heist EP}, \textbf{Moe LA}. 2015.
Bacterial community structure and dynamics during corn-based bioethanol
fermentation. Microbial Ecology \textbf{71}:409--421.
doi:\href{https://doi.org/10.1007/s00248-015-0673-9}{10.1007/s00248-015-0673-9}.}

\leavevmode\vadjust pre{\hypertarget{ref-Baxter2016}{}}%
\CSLLeftMargin{4. }%
\CSLRightInline{\textbf{Baxter NT}, \textbf{Ruffin MT}, \textbf{Rogers
MAM}, \textbf{Schloss PD}. 2016. Microbiota-based model improves the
sensitivity of fecal immunochemical test for detecting colonic lesions.
Genome Medicine \textbf{8}.
doi:\href{https://doi.org/10.1186/s13073-016-0290-3}{10.1186/s13073-016-0290-3}.}

\leavevmode\vadjust pre{\hypertarget{ref-Beall2015}{}}%
\CSLLeftMargin{5. }%
\CSLRightInline{\textbf{Beall BFN}, \textbf{Twiss MR}, \textbf{Smith
DE}, \textbf{Oyserman BO}, \textbf{Rozmarynowycz MJ}, \textbf{Binding
CE}, \textbf{Bourbonniere RA}, \textbf{Bullerjahn GS}, \textbf{Palmer
ME}, \textbf{Reavie ED}, \textbf{Waters LMK}, \textbf{Woityra LWC},
\textbf{McKay RML}. 2015. Ice cover extent drives phytoplankton and
bacterial community structure in a large north-temperate lake:
Implications for a warming climate. Environmental Microbiology
\textbf{18}:1704--1719.
doi:\href{https://doi.org/10.1111/1462-2920.12819}{10.1111/1462-2920.12819}.}

\leavevmode\vadjust pre{\hypertarget{ref-Henson2016}{}}%
\CSLLeftMargin{6. }%
\CSLRightInline{\textbf{Henson MW}, \textbf{Pitre DM}, \textbf{Weckhorst
JL}, \textbf{Lanclos VC}, \textbf{Webber AT}, \textbf{Thrash JC}. 2016.
Artificial seawater media facilitate cultivating members of the
microbial majority from the gulf of mexico. {mSphere} \textbf{1}.
doi:\href{https://doi.org/10.1128/msphere.00028-16}{10.1128/msphere.00028-16}.}

\leavevmode\vadjust pre{\hypertarget{ref-Baxter2014}{}}%
\CSLLeftMargin{7. }%
\CSLRightInline{\textbf{Baxter NT}, \textbf{Wan JJ}, \textbf{Schubert
AM}, \textbf{Jenior ML}, \textbf{Myers P}, \textbf{Schloss PD}. 2014.
Intra- and interindividual variations mask interspecies variation in the
microbiota of sympatric peromyscus populations. Applied and
Environmental Microbiology \textbf{81}:396--404.
doi:\href{https://doi.org/10.1128/aem.02303-14}{10.1128/aem.02303-14}.}

\leavevmode\vadjust pre{\hypertarget{ref-LevyBooth2018}{}}%
\CSLLeftMargin{8. }%
\CSLRightInline{\textbf{Levy-Booth DJ}, \textbf{Giesbrecht IJW},
\textbf{Kellogg CTE}, \textbf{Heger TJ}, \textbf{D'Amore DV},
\textbf{Keeling PJ}, \textbf{Hallam SJ}, \textbf{Mohn WW}. 2018.
Seasonal and ecohydrological regulation of active microbial populations
involved in {DOC}, {CO}2, and {CH}4 fluxes in temperate rainforest soil.
The {ISME} Journal \textbf{13}:950--963.
doi:\href{https://doi.org/10.1038/s41396-018-0334-3}{10.1038/s41396-018-0334-3}.}

\leavevmode\vadjust pre{\hypertarget{ref-Edwards2015}{}}%
\CSLLeftMargin{9. }%
\CSLRightInline{\textbf{Edwards J}, \textbf{Johnson C},
\textbf{Santos-Medellín C}, \textbf{Lurie E}, \textbf{Podishetty NK},
\textbf{Bhatnagar S}, \textbf{Eisen JA}, \textbf{Sundaresan V}. 2015.
Structure, variation, and assembly of the root-associated microbiomes of
rice. Proceedings of the National Academy of Sciences
\textbf{112}:E911--E920.
doi:\href{https://doi.org/10.1073/pnas.1414592112}{10.1073/pnas.1414592112}.}

\leavevmode\vadjust pre{\hypertarget{ref-Ettinger2017}{}}%
\CSLLeftMargin{10. }%
\CSLRightInline{\textbf{Ettinger CL}, \textbf{Williams SL},
\textbf{Abbott JM}, \textbf{Stachowicz JJ}, \textbf{Eisen JA}. 2017.
Microbiome succession during ammonification in eelgrass bed sediments.
{PeerJ} \textbf{5}:e3674.
doi:\href{https://doi.org/10.7717/peerj.3674}{10.7717/peerj.3674}.}

\leavevmode\vadjust pre{\hypertarget{ref-Graw2018}{}}%
\CSLLeftMargin{11. }%
\CSLRightInline{\textbf{Graw MF}, \textbf{DAngelo G}, \textbf{Borchers
M}, \textbf{Thurber AR}, \textbf{Johnson JE}, \textbf{Zhang C},
\textbf{Liu H}, \textbf{Colwell FS}. 2018. Energy gradients structure
microbial communities across sediment horizons in deep marine sediments
of the south china sea. Frontiers in Microbiology \textbf{9}.
doi:\href{https://doi.org/10.3389/fmicb.2018.00729}{10.3389/fmicb.2018.00729}.}

\leavevmode\vadjust pre{\hypertarget{ref-Johnston2016}{}}%
\CSLLeftMargin{12. }%
\CSLRightInline{\textbf{Johnston ER}, \textbf{Rodriguez-R LM},
\textbf{Luo C}, \textbf{Yuan MM}, \textbf{Wu L}, \textbf{He Z},
\textbf{Schuur EAG}, \textbf{Luo Y}, \textbf{Tiedje JM}, \textbf{Zhou
J}, \textbf{Konstantinidis KT}. 2016. Metagenomics reveals pervasive
bacterial populations and reduced community diversity across the alaska
tundra ecosystem. Frontiers in Microbiology \textbf{7}.
doi:\href{https://doi.org/10.3389/fmicb.2016.00579}{10.3389/fmicb.2016.00579}.}

\leavevmode\vadjust pre{\hypertarget{ref-Hassell2018}{}}%
\CSLLeftMargin{13. }%
\CSLRightInline{\textbf{Hassell N}, \textbf{Tinker KA}, \textbf{Moore
T}, \textbf{Ottesen EA}. 2018. Temporal and spatial dynamics in
microbial community composition within a temperate stream network.
Environmental Microbiology \textbf{20}:3560--3572.
doi:\href{https://doi.org/10.1111/1462-2920.14311}{10.1111/1462-2920.14311}.}

\end{CSLReferences}

\bibliography{ref}
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}

\textbf{Table 1. Summary of studies used in the analysis.} For all
studies, the number of sequences used from each study was rarefied to
the smallest sample size. A graphical represenation of the distribution
of sample sizes for each study and the samples that were removed from
each study are provided in Figure S1.

\small

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1367}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0664}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1914}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1953}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2031}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2070}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Study\nobreakspace{}(Ref)}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\textbf{Samples}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\makecell[c]{\textbf{Total}\\\textbf{sequences}}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\makecell[c]{\textbf{Median}\\\textbf{sequences}}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\makecell[c]{\textbf{Range of}\\\textbf{sequences}}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\makecell[c]{\textbf{SRA study}\\\textbf{accession}}
\end{minipage} \\
\midrule()
\endhead
Bioethanol~(3) & 95 & 3,970,972 & 16,014 & 3,690\Hyphdash*356,027 &
SRP055545 \\
Human~(4) & 490 & 20,828,275 & 32,452 & 10,439\Hyphdash*422,904 &
SRP062005 \\
Lake~(5) & 52 & 3,145,486 & 69,205 & 15,135\Hyphdash*110,993 &
SRP050963 \\
Marine~(6) & 7 & 1,484,068 & 213,091 & 132,895\Hyphdash*256,758 &
SRP068101 \\
Mice~(1) & 348 & 2,785,641 & 6,426 & 1,804\Hyphdash*30,311 &
SRP192323 \\
Peromyscus~(7) & 111 & 1,545,288 & 12,393 & 4,454\Hyphdash*33,502 &
SRP044050 \\
Rainforest~(8) & 69 & 936,666 & 11,464 & 4,880\Hyphdash*37,403 &
ERP023747 \\
Rice~(9) & 490 & 22,623,937 & 43,399 & 2,777\Hyphdash*192,200 &
SRP044745 \\
Seagrass~(10) & 286 & 4,135,440 & 13,538 & 1,830\Hyphdash*45,076 &
SRP092441 \\
Sediment~(11) & 58 & 1,151,389 & 17,606 & 7,686\Hyphdash*67,763 &
SRP097192 \\
Soil~(12) & 18 & 932,563 & 50,487 & 46,622\Hyphdash*58,935 &
ERP012016 \\
Stream~(13) & 201 & 21,017,610 & 90,621 & 8,931\Hyphdash*394,419 &
SRP075852 \\
\bottomrule()
\end{longtable}

\normalsize
